A Systematic Literature Review on Extensions to Role-Based Access Control

### 2014-02-08 paper/sections/analysis.tex
updates - 1. Reorganization results section 2. update Figure 2
Updating data display tables.

@@

During the paper reading phase, we identify motivations of the RBAC extension models by category. We identify motivations by looking for issues regarding why the RBAC reference model was inadequate and how the authors addressed the issues. For example, the papers in the Spatial Category argue that the RBAC reference model does not support spatial constraints, which are used in determining users' access decisions based on spatial dimensions (i.e., location). To address this issue, the authors developed the RBAC extension models by incorporating additional spatial entities and their relations with existing entities of the RBAC Reference Model. We describe motivations behind the RBAC extension models by category as follows:

@@

\begin{table}
\centering
\caption{Implementation types found and the count of primary sources}
\begin{tabular}{ | c | c | }
\cline{1-2}

\textbf{Implementation Type} & \textbf{Paper Count} \\ \cline{1-2}
Enterprise Implementation & 4 \\ \cline{1-2}
Prototype Implementation & 4 \\ \cline{1-2}
No Implementation & 20 \\

\cline{1-2}
\end{tabular}
\label{tab:implementations}
\end{table}

@@

@@

\begin{itemize}
\setlength{\itemsep}{0.25pt}
\item Time-based Performance \cite{ni2010privacy}, \cite{aich09:role}
\item Complexity analysis \cite{bao08:role}, \cite{zhang06:collaborative}, \cite{chen08:spatio-temporal}, \cite{aich09:role}
\item Comparison to standard RBAC \cite{bao08:role}, \cite{zou2009crbac}, \cite{zhang06:collaborative}, \cite{zhao2008flexible}, \cite{ray07:spatio}
\item Mathematical modeling \cite{damiani2007geo}, \cite{hansen2003spatial}, \cite{aich07:STARBAC}, \cite{chen08:spatio-temporal}, \cite{joshi05:generalized}
\item Example scenarios of the model in action \cite{alam06:constraint}, \cite{tzelepi01:flexible}, \cite{cholewka00:acontext-sensitive}, \cite{huang06:pervasive}, \cite{bao08:role}, \cite{jian2008extended}, \cite{yamazaki04:designing}, \cite{zou2009crbac}, \cite{ray07:spatio}, \cite{samuel07:spatio-temporal}, \cite{joshi05:generalized}, \cite{yao2008task}, \cite{zhou2007team}, \cite{oh2003task}
%\item Experimental analysis of the model
\item Case study of the model in practice \cite{motta03:contextual}
\end{itemize}

@@

@@

@@

Business needs have historically driven RBAC research and development.  The primary mode of evaluation for
model extensions has been the presentation of business scenarios in various domains and how the model
uniquely handles those particular scenarios.  Thus, looking for trends in the domains used in the example
scenarios might serve to illuminate a trend worth further examination into the reason for the explosion of
RBAC extensions.  We identified domains presented within the primary sources by looking for example
scenarios cast within a particular domain or mention of domain requirements within the body of the paper.
We found that the domains mentioned and their associated sources are:

\begin{itemize}
\setlength{\itemsep}{0.25pt}
\item Medical domain \cite{alam06:constraint}, \cite{tzelepi01:flexible}, \cite{motta03:contextual}, \cite{ni2010privacy}, \cite{damiani2007geo}, \cite{hansen2003spatial}, \cite{samuel07:spatio-temporal}, \cite{aich09:role}, \cite{zhou2007team}
\item Pervasive computing environments \cite{huang06:pervasive}, \cite{chen08:spatio-temporal}, \cite{ray07:spatio}
\item Web applications \cite{masoumzadeh2008purbac}
\item Mobile computing \cite{thein2011leveraging}, \cite{zou2009crbac}, \cite{chandran05:llt}, \cite{ray07:spatio}, \cite{aich09:role}
\item Large-scale organizations with many sub-departments \cite{yamazaki04:designing}, \cite{han08:extended}, \cite{yao2008task}
\item Enterprise, organization workflows \cite{cholewka00:acontext-sensitive}, \cite{bao08:role}, \cite{zhang06:collaborative}, \cite{oh2003task}, \cite{joshi05:generalized}
\end{itemize}

@@

>>>>

%\subsection{Results}

During the paper reading phase, we identify motivations of the RBAC extension models by category. We identify motivations by looking for issues regarding why the RBAC reference model was inadequate and how the authors addressed the issues. For example, the papers in the Spatial Category argue that the RBAC reference model does not support spatial constraints, which are used in determining users' access decisions based on spatial dimensions (i.e., location). To address this issue, the authors developed the RBAC extension models by incorporating additional spatial entities and their relations with existing entities of the RBAC Reference Model.

@@

We describe motivations behind the RBAC extension models by category as follows:

@@

@@

\begin{table}
\centering
\caption{Implementation types found and the count of primary sources}
\vspace{0.1 in}
\begin{tabular*}{.9\linewidth}{| l | p{.5\linewidth} | r | }
\cline{1-3}

\textbf{Implementation Type} & \textbf{Papers} & \textbf{Count} \\ \cline{1-3}
Enterprise
&
\cite{motta03:contextual}
\cite{aich09:role}
\cite{yao2008task}
&
3 \\ \cline{1-3}
Prototype
& 
\cite{cholewka00:acontext-sensitive}
\cite{huang06:pervasive}
\cite{bao08:role}
\cite{zhou07:network}
&
4 \\ \cline{1-3}
None
&
\cite{alam06:constraint}
\cite{tzelepi01:flexible}
\cite{yamazaki04:designing}
\cite{jian2008extended}
\cite{thein2011leveraging}
\cite{zou2009crbac}
\cite{haibo05:context}
\cite{zhang06:collaborative}
\cite{ni2010privacy}
\cite{masoumzadeh2008purbac}
\cite{damiani2007geo}
\cite{ray2006lrbac}
\cite{hansen2003spatial}
\cite{aich07:STARBAC}
\cite{chen08:spatio-temporal}
\cite{samuel07:spatio-temporal}
\cite{chandran05:llt}
\cite{ray07:spatio}
\cite{oh2003task}
\cite{joshi05:generalized}
&
20 \\

\cline{1-3}
\end{tabular*}
\label{tab:implementations}
\end{table}

@@

@@

\begin{table}
\centering
\caption{Evaluation types by primary source}
\vspace{0.1 in}
\begin{tabular*}{.9\linewidth}{| p{.45\linewidth} | p{.3\linewidth} | r | }
\cline{1-3}

\textbf{Evaluation Type} & \textbf{Papers} & \textbf{Count} \\ \cline{1-3}
Time-Based Performance
&
\cite{ni2010privacy}
\cite{aich09:role}
&
2 \\ \cline{1-3}
Complexity Analysis
&
\cite{bao08:role}
\cite{zhang06:collaborative}
\cite{chen08:spatio-temporal}
\cite{aich09:role}
&
4 \\ \cline{1-3}
Comparison to Standard RBAC
& 
\cite{bao08:role}
\cite{zou2009crbac}
\cite{zhang06:collaborative}
\cite{zhao2008flexible}
\cite{ray07:spatio}
&
5 \\ \cline{1-3}
Mathematical Modeling
& 
\cite{damiani2007geo}
\cite{hansen2003spatial}
\cite{aich07:STARBAC}
\cite{chen08:spatio-temporal}
\cite{joshi05:generalized}
&
5 \\ \cline{1-3}
Example Scenarios of the Model in Action
& 
\cite{alam06:constraint}
\cite{tzelepi01:flexible}
\cite{cholewka00:acontext-sensitive}
\cite{huang06:pervasive}
\cite{bao08:role}
\cite{jian2008extended}
\cite{yamazaki04:designing}
\cite{zou2009crbac}
\cite{ray07:spatio}
\cite{samuel07:spatio-temporal}
\cite{joshi05:generalized}
\cite{yao2008task}
\cite{zhou2007team}
\cite{oh2003task}
&
14 \\ \cline{1-3}
Case Study of the Model in Practice
&
\cite{motta03:contextual}
&
1 \\ \cline{1-3}

\end{tabular*}
\label{tab:implementations}
\end{table}

@@

%\subsection{Results}

Business needs have historically driven RBAC research and development.  The primary mode of evaluation for
model extensions has been the presentation of business scenarios in various domains and how the model
uniquely handles those particular scenarios.  Thus, looking for trends in the domains used in the example
scenarios might serve to illuminate a trend worth further examination into the reason for the explosion of
RBAC extensions.  We identified domains presented within the primary sources by looking for example
scenarios cast within a particular domain or mention of domain requirements within the body of the paper.

@@

We found that the domains mentioned and their associated sources are:

\begin{table}
\centering
\caption{Domains by primary source}
\vspace{0.1 in}
\begin{tabular*}{.9\linewidth}{| p{.45\linewidth} | p{.3\linewidth} | r | }
\cline{1-3}

\textbf{Domain} & \textbf{Papers} & \textbf{Count} \\ \cline{1-3}
Medical Domain
&
\cite{alam06:constraint}
\cite{tzelepi01:flexible}
\cite{motta03:contextual}
\cite{ni2010privacy}
\cite{damiani2007geo}
\cite{hansen2003spatial}
\cite{samuel07:spatio-temporal}
\cite{aich09:role}
\cite{zhou2007team}
&
9 \\ \cline{1-3}
Pervasive Computing Environments
& 
\cite{huang06:pervasive}
\cite{chen08:spatio-temporal}
\cite{ray07:spatio}
&
3 \\ \cline{1-3}
Web Applications
& 
\cite{masoumzadeh2008purbac}
&
1 \\ \cline{1-3}
Mobile Computing
& 
\cite{thein2011leveraging}
\cite{zou2009crbac}
\cite{chandran05:llt}
\cite{ray07:spatio}
\cite{aich09:role}
&
5 \\ \cline{1-3}
Organizations with Many Sub-departments
& 
\cite{yamazaki04:designing}
\cite{jian2008extended}
\cite{yao2008task}
&
3 \\ \cline{1-3}
Enterprise Workflows
&
\cite{cholewka00:acontext-sensitive}
\cite{bao08:role}
\cite{zhang06:collaborative}
\cite{oh2003task}
\cite{joshi05:generalized}
&
5 \\ \cline{1-3}

\end{tabular*}
\label{tab:implementations}
\end{table}

@@

During the paper reading phase, we identified commonalities or generalizations within the primary sources 
by looking for formal modeling of extending RBAC extension models.

### 2014-02-08 paper/sections/appendix.tex
Committing updates based on comments from reading group.
Adding greybox. Adding results subsections for each RQ. Fixing last table captions to be on top. Edits to categories.
Removing duplicate paper and updating total counts.
Updating data display tables.

\begin{sidewaystable}
\centering
\vline
\begin{tabularx}{\linewidth}{ | X | X | X | X | }
\cline{1-4}

@@

\multicolumn{1}{ c }{}  &
\multicolumn{3}{ |l| }{An extended RBAC model based on granular logic, 2008 \cite{han08:extended}} \\ \cline{2-4}

@@

\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{ Res. }}}} &
\multicolumn{2}{ |l| }{A Flexible Role- and Resource-Based Access Control Model, 2008 \cite{zhao2008flexible}} \\[2.5ex]  \cline{2-4}

@@

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{}  &
\multicolumn{1}{ |l| }{LRBAC: A location-aware role-based access control model, 2006 \cite{ray07:spatio}} \\[0.25ex] \cline{4-4}

@@

\cline{1-4}
\end{tabularx}
\vline
\caption{Primary sources grouped by categorization}
\label{tab:categorization}
\end{sidewaystable}

>>>>

\begin{sidewaystable}
\centering
\caption{Primary sources grouped by categorization}
\vline
\begin{tabularx}{\linewidth}{ | X | X | X | X | }
\cline{1-4}

@@

@@

@@

\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{}  &
\multicolumn{1}{ |l| }{LRBAC: A location-aware role-based access control model, 2006 \cite{ray2006lrbac}} \\[0.25ex] \cline{4-4}

@@

\cline{1-4}
\end{tabularx}
\vline
\label{tab:categorization}
\end{sidewaystable}
